\thetitle

### 2021-03-02 writing/challenges.tex
pass over challenges section

The {\em first challenge\/} stems from the interplay between blame and the
run-time checks of a gradual type systems. When a check fails the Natural
and Transient semantics assign blame and discard  the information in the
exception from the run-time check. Nevertheless, the exception
information may be equally or even more useful to the rational programmer
than blame. If this is the case blame may not play a critical role for the
rational programmer.\footnote{Indeed, precisely because they do not
account for confounding factors, ~\citet{lksfd-popl-2020} cannot draw any
conclusions about the pragmatics of blame despite advertisements for the
opposite. Their experiment can only conclude that so-called blame-shifting
works but not that this is because of blame.}

One way to isolate blame from this confounding factor is to {\em equip the
rational programmer with modes \/} that represent the different types of
information the rational programmer takes into account when debugging a
scenario. Specifically, the rational programmer comes with a blame and
an exception mode. Only if the former succeeds debugging a program while
the latter fails, it is safe to conclude that blame is indeed necessary
for the rational programmer. Put differently, the exception mode serves as
the {\em baseline\/} for blame's value; if the programmer in this mode
``beats'' the  blame one, a blame assignment mechanism might be useless.

Besides the run-time checks, pure luck might also amplify the usefulness
of blame. Thus a proper evaluation of blame needs to factor out any random
effects. A random rational programmer mode offers the necessary baseline
using the same ingredients as for exceptions.

@@

With modes, different strategy map to different ways the rational
programmer  interprets blame.  For instance, one Transient mode may pick
the oldest element of a blame set as the next component to add types
because it corresponds to the earliest point in the program's evaluation
that can discover an impedance mismatch.  Another mode may opt to treat
the sequence as a stack and add types to its newest element.  When both
modes are equally successful in locating an impedance mismatch, measuring
the rational programmer's effort may answer which mode is the most
effective.

>>>>

The {\em first challenge\/} stems from the interplay between blame and the
run-time checks of a gradual type systems. When a check fails the Natural
and Transient semantics assign blame and discard  the information in the
exception from the run-time check. Nevertheless, the exception
information may be equally or even more useful to the rational programmer
than blame. If this is the case blame may not play a critical role for the
rational programmer. Indeed, precisely because they do not
account for confounding factors, ~\citet{lksfd-popl-2020} cannot draw any
conclusions about the pragmatics of blame despite advertisements for the
opposite. Their experiment can only conclude that so-called blame-shifting
works but not that this is because of blame.

One way to isolate blame from this confounding factor is to {\em equip the
rational programmer with modes \/} that represent the different types of
information the rational programmer takes into account when debugging a
scenario. Specifically, the rational programmer comes with a blame mode and
an exception mode. Only if the former succeeds debugging a program while
the latter fails, it is safe to conclude that blame is indeed necessary
for the rational programmer. Put differently, the exception mode serves as
the {\em baseline\/} for blame's value; if the programmer in this mode
``beats'' the  blame one, a blame assignment mechanism might be useless.

Besides the run-time checks, pure luck might also amplify the usefulness
of blame. Thus a proper evaluation of blame needs to factor out any random
effects. A random rational programmer mode provides the necessary baseline
using the same ingredients as for exceptions.

@@

With modes, different blame strategies map to different ways the rational
programmer  interprets blame.  For instance, one Transient mode may pick
the oldest element of a blame set as the next component to add types
because it corresponds to the earliest point in the program's evaluation
that can discover an impedance mismatch.  Another mode may opt to treat
the sequence as a stack and add types to its newest element.  When both
modes are equally successful in locating an impedance mismatch, measuring
the rational programmer's effort may answer which mode is the most
effective.
